---
title: "Hurricane Public Assistance"
author: "Jingwen Xu"
date: "2020/11/9"
output: html_document
---

```{r setup, include=FALSE}
library(readr)
library(magrittr)
library(dplyr)
library(ggplot2)
library(tmap)
library(sf)
library(maps)
library(tidyr)
library(kableExtra)
knitr::opts_chunk$set(echo = FALSE, 
               warning = FALSE,
               message = FALSE)
```


# Data cleaning and organization

We only consider about the hurricane disaster during 2009-2018. So before mapping, I need to simply filter and clean the data.

```{r}
disaster <- read.csv("PublicAssistanceFundedProjectsDetails.csv")

## Filter the data about hurricane
hurricane <- disaster %>% filter(incidentType=="Hurricane")
## unique(hurricane$incidentType)
hurricane %<>% select(-3)

## We only study the hurricane declared during 2009-2018.
hurricane$year <- substr(hurricane$declarationDate,1,4)
hurricane %<>% filter(year=="2009"|year=="2010"|year=="2011"|year=="2012"|year=="2013"|year=="2014"|year=="2015"|year=="2016"|year=="2017"|year=="2018")

## summary(hurricane$totalObligated-hurricane$federalShareObligated) These two columns are totally the same.
## summary(hurricane)
## We found that there are negative values in "projectAmount", "federalShareObligated", "totalObligated" which are not reasonable that should be deleted.
hurricane %<>% filter(projectAmount>=0&federalShareObligated>=0)
kable(head(hurricane)) %>% kable_styling(font_size=12)
```

# Mapping

I map the data using two methods -- ggplot2 and tmap. And there are three data information that I want to show on the map -- the count of hurricane for each county, the total project amount and total federal obligated per county.

## Mapping with tmap

```{r}
hurricane$county <- tolower(hurricane$county)
hurricane$state <- tolower(hurricane$state)
Map <- st_as_sf(map('county',plot=F,fill=T))
Map%<>%separate(ID,c("state","county"),sep=",")
hurr_count <- hurricane %>% group_by(county,state) %>% summarise(Count=n())
hurr_count_t <- left_join(Map,hurr_count,by=c("county","state"))
hurr_count_t$Count[is.na(hurr_count_t$Count)]=0.01
hurr_count_t$Count <- cut(hurr_count_t$Count,breaks=c(-1,1,8,24,75,85,1615),labels=c("[0,1]","(1,8]","(8,24]","(24,75]","(75,85]","(85,1615]"),ordered_result=TRUE)
tm_shape(hurr_count_t,title="The count of hurricane per county")+
  tm_polygons("Count",palette="Blues",title="Count of hurricane")+tm_layout(main.title='The count of hurricane per county')
```

```{r}
assist <- hurricane %>% group_by(county,state) %>% summarise(project_t=sum(projectAmount))
assist_t <- left_join(Map,assist,by=c("county","state"))
assist_t$project_t[is.na(assist_t$project_t)]=0.01
assist_t$project_t <- cut(assist_t$project_t,breaks=c(-1,1.356e+03,1.701e+05,7.970e+05,4.651e+06,3.120e+07,9.801e+09),labels=c("[0,1.356e+03]","(1.356e+03,1.701e+05]","(1.701e+05,7.970e+05]","(7.970e+05,4.651e+06]","(4.651e+06,3.120e+07]","(3.120e+07,9.801e+09]"),ordered_result = T)
tm_shape(assist_t,title="The total public assistance cost per county")+tm_polygons("project_t",palette="GnBu",title="Public assistance cost")+tm_layout(main.title='The total public assistance cost per county')
```

```{r}
federal <- hurricane %>% group_by(county,state) %>% summarise(federal_total=sum(federalShareObligated))
federal_t <- left_join(Map,federal,by=c("county","state"))
federal_t$federal_total[is.na(federal_t$federal_total)]=0.01
federal_t$federal_total <- cut(federal_t$federal_total,breaks=c(-1,1.017e+03,1.340e+05,6.223e+05,3.638e+06,2.751e+07,8.900e+09),labels=c("[0,1.017e+03]","(1.017e+03,1.340e+05]","(1.340e+05,6.223e+05]","(6.223e+05,3.638e+06]","(3.638e+06,2.751e+07]","(2.751e+07,8.809e+09]"),ordered_result = T)
tm_shape(federal_t,title="The total Federal Share Obligated per county")+tm_polygons("federal_total",palette="Oranges",title="Federal share obligated")+tm_layout(main.title='The total Federal Share Obligated per county')
```

## Mapping with ggplot

```{r}
county <- map_data("county")
state <- map_data("state")
ggplot() + ggtitle("The count of hurricane per county")+
  geom_polygon(data=county, aes(x=long, y=lat, group=group),
               color="grey", fill="white", size = .2, alpha=0.9) +
  geom_polygon(data=state, aes(x=long, y=lat, group=group),
               color="black", fill="white", size = .2, alpha = .1)+
  geom_sf(data=hurr_count_t,mapping=aes(fill = Count))+
  scale_fill_brewer(name="Count of hurricane")
```

```{r}
ggplot() + ggtitle("The total public assistance cost per county")+
  geom_polygon(data=county, aes(x=long, y=lat, group=group),
               color="grey", fill="white", size = .2, alpha=0.9) +
  geom_polygon(data=state, aes(x=long, y=lat, group=group),
               color="black", fill="white", size = .2, alpha = .1)+
  geom_sf(data=assist_t,mapping=aes(fill = project_t))+
  scale_fill_brewer(name="Public assistance cost",type="seq",palette=4)
```

```{r}
ggplot() + ggtitle("The total Federal Share Obligated per county")+
  geom_polygon(data=county, aes(x=long, y=lat, group=group),
               color="grey", fill="white", size = .2, alpha=0.9) +
  geom_polygon(data=state, aes(x=long, y=lat, group=group),
               color="black", fill="white", size = .2, alpha = .1)+
  geom_sf(data=federal_t,mapping=aes(fill = federal_total))+
  scale_fill_brewer(name="Federal share obligated",type="seq",palette=7)
```

# Conclusion

From the mappings for three data information, we can find that the distribution of count of hurricane, total project amount and total federal obligated are similar which is reasonable. The county with more hurricane need more public assistance and federal obligated.
